---
title: "HW07 - Digit Recognition"
author: "Leonard Armstrong"
date: "2/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CLEAN START

Create a clean start anytime the script is executed from the top.

```{r clean_start}

##-------------------------------------------------------------------------------------------------
## CLEAN START

# Remove any non-default packages
default_packages <- union(getOption("defaultPackages"), "base")
loaded_packages <- .packages()
extra_packages <- setdiff(x = loaded_packages, y = default_packages)
invisible(
  sapply(
    X = extra_packages, 
    FUN = function(x) detach(name = paste("package:", x, sep=""), character.only = TRUE)))

# Remove all variables and functions from the global environment
rm(list = ls())
```

# LOAD LIBRARIES

Load all libraries required for this script.

Library      | Purpose
------------ | ---------------------------------------------------------------------
assertthat   | Create and manage assertions for ensuring proper states in the code.
class        | Classification routines (knn).
e1071        | Classification model functions. (naiveBayes and svm)
ggplot2      | Graphing and plotting functions.
gridExtra    | Grid layout for graphics.
here         | Working directory utilities.
randomForest | Random forest functions.
RColorBrewer | Color palette definitions.
reshape      | Melt function.
scales       | ggplot2 axis scaling functions.
stringi      | String management functions.

```{r load_libraries}

##-------------------------------------------------------------------------------------------------
## LOAD LIBRARIES
##
## This section loads all required libraries.
##
## Library      | Purpose
## ------------ | ---------------------------------------------------------------------
## assertthat   | Create and manage assertions for ensuring proper states in the code.
## class        | Classification routines (knn).
## e1071        | Naive Bayes functionality.
## ggplot2      | Graphing and plotting functions.
## gridExtra    | Grid layout for graphics.
## here         | Working directory utilities.
## randomForest | Random forest functions.
## RColorBrewer | Color palette definitions.
## reshape      | Melt function.
## scales       | ggplot2 axis scaling functions.
## stringi      | String management functions.

library(assertthat)
library(class)
library(e1071)
library(ggplot2)
library(gridExtra)
library(here)
library(randomForest)
library(RColorBrewer)
library(reshape)
library(scales)
library(stringi)
```

# DEFINE KEY CONSTANTS

```{r define_constants}

##-------------------------------------------------------------------------------------------------
## DEFINE KEY CONSTANTS AND OTHER DATA TO BE USED THROUGHOUT THE CODE

# Define assertion error messages
errmsg_trainfilenotexist <- "The specified training file does not exist."
errmsg_testfilenotexist <- "The specified test file does not exist."

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 

# Define the primary random seed (re)used throughout the script.
random_seed <- 100163

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 

# Define number of folds for k-folds testing.
k_folds <- 10

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 

# Define observation sizes
image_height <- 28
image_width  <- 28
image_size   <- image_height * image_width
record_size  <- image_size + 1              # Add one for the label field. 

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 

# Define the relevant directories, file names, and file paths.
cwd <- here::here()
data_subdir <- "../data"

# Define the training data file.
training_filename <- "Kaggle-digit-train-sample-small-1400.csv"
training_fullfilepath <- file.path(cwd, data_subdir, training_filename)
assertthat::assert_that(file.exists(training_fullfilepath), msg = errmsg_trainfilenotexist)

# Define the test data file.
test_filename <- "Kaggle-digit-test-sample1000.csv"
test_fullfilepath <- file.path(cwd, data_subdir, test_filename)
assertthat::assert_that(file.exists(training_fullfilepath), msg = errmsg_trainfilenotexist)
```

```{r remove_vars1}

# Remove no-longer-needed variables.
rm(training_filename, test_filename)
```

# READ DATA

## Training Data

### Read training data

```{r read_data}

##-------------------------------------------------------------------------------------------------
## READ PROVIDED TRAINING DATA SET

# Set the random seed
set.seed(100163)

# Define training data datatypes as numerics. Label is a factor and pixels are numeric.
numeric_coltypes <- c("factor", rep(x = "numeric", times = image_size))  

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
# Review the training data set.

# Read the training data and record the number of observations.
train_num <- read.csv(
  file = training_fullfilepath, header = TRUE, colClasses = numeric_coltypes, na.strings = "")

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
# Define key k-fold test sizes based on the training data size.

# Get number of rows in training data set
train_nrows <- nrow(train_num)

# Now that we know the volume of data, define the size of a fold.
fold_size <- round(train_nrows/k_folds)

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
# Review key values in the data set.

# Verify the minimum and maximum pixel values.
pixel_min <- min(train_num[, -1])
pixel_max <- max(train_num[, -1])
cat("The minumim training data value is", pixel_min, "\n")
cat("The maximum training data value is", pixel_max, "\n")

```

### Shuffle the base training data

```{r initialize_analysis_dataset}

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
## Shuffle the original data set before creating any subsequent data sets.

set.seed(seed = random_seed)

# Shuffle the base data set
train_num <- train_num[sample(train_nrows), ]

```

### Create binary numeric data set

```{r create_binary_numeric_training_data}

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
# Create a binary_numeric training data set.

train_binary <- data.frame(label = train_num$label, sign(train_num[, -1] > 32))

```

### Clean the training data

```{r clean_training_data}

##-------------------------------------------------------------------------------------------------
## REVIEW AND CLEAN (IF NEEDED) PROVIDED DATA SETS
##
## In this section, the data sets will be reviewed for "cleanliness" and cleaned if needed.

# Review NAs in the data
cat("There are NAs in the decimal training data:", any(is.na(train_num)), "\n")
cat("There are NAs in the binary training data:", any(is.na(train_binary)), "\n")

```

## Test Data
```{r read_test_data}

##-------------------------------------------------------------------------------------------------
## READ AND CLEAN PROVIDED TEST DATA SET

# Read the test data.
test_num <- read.csv(
  file = test_fullfilepath, header = TRUE, colClasses = numeric_coltypes, na.strings = "")
nas_in_test <- any(is.na(test_num))
cat("The test data", ifelse(nas_in_test, "DOES", "does NOT"), "contain NA values.\n")
if (nas_in_test) {
  cat("Removing NA records...\n", sep = "")
  test_num <- na.omit(test_num)
  nas_in_test <- any(is.na(test_num))
  cat("The test data", ifelse(nas_in_test, "DOES", "does NOT"), "contain NA values.\n")
}

# Create a binary test data set
test_binary <- data.frame(label = test_num$label, sign(test_num[, -1] > 32))
```

# DATA EDA SUMMARIES

```{r summarize_training_data, eval=FALSE}

##-------------------------------------------------------------------------------------------------
## EXPLORATORY DATA ANALYSIS - DATA SUMMARIES
##
## This section, consists of numeric exploratory analysis of the data.

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
# Summarize the labels

# Show basic statistic summary of the training data.
cat("Summary of the training data:\n", sep="")
summary(as.numeric(levels(train_num$label)))
cat("Standard deviation of the training data: ", sd(as.numeric(train_num$label)), "\n", sep="")

# Review the means and standard deviations of each pixel and determine which pixels have zeros
# for all training cases. 
all_means <- apply(X=train_num[, -1], MARGIN = 2, FUN = mean)
all_zeros <- all_means[all_means == 0]
cat(NROW(all_zeros), " pixels have zeros for every record.\n", sep="")

# Remove unnecessary data
rm (all_means, all_sds, all_zeros)

```

```{r review_pixel_distribution}

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
# Review the distribution of values across all possible pixels. 

# Turn the tabluar data into a longer data model.
train_long <- melt(data = train_num)

# Change range from 0-255 to 1-256 to avoind log(0) errors.
train_long$value <- train_long$value + 1

# Determine the distribution of values for all pixels
distrib <- table(train_long$value)
distrib_df <- data.frame(distrib, row.names = names(distrib))
names(distrib_df) <- c("pixval", "frequency")

# Create a historgram of the distribution of pixel values.
# Set y-axis to log scale to magnify lower values.
bar_breaks <- c(1, 10, 100, 1000, 10000, 100000, 1000000)
gpixbar <- ggplot(data = distrib_df, mapping = aes(x = pixval, y = frequency, fill = frequency)) +
  geom_col() + 
  labs(
    title = "Frequency of Possible Pixel Values Across All 1,097,600 Training Set Pixels",
    subtitle = "1400 training set images × 784 pixels/image = 1,097,600 training set pixels",
    x = "Pixel Value",
    y = "Frequency") +
  scale_y_log10(labels = scales::comma, breaks = bar_breaks) + 
  scale_fill_gradient(low = "lightblue", high = "navy", trans = "log", breaks = bar_breaks) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  theme(legend.position = "none")
gpixbar
```

# VISUAL EDA

## Create a bar chart of digit frequencies.

```{r visual_eda_digit_barplot}

##-------------------------------------------------------------------------------------------------
## EXPLORATORY DATA ANALYSIS - DATA VISUALIZATIONS
##
## This section, consists of visual exploratory analysis of the data.

# Create a support dataframe of labels, label counts and label percentages
lcount <- table(train_num$label)
lpct <- c(lcount/sum(lcount), use.names = FALSE)
pctlab <- sprintf("%5.2f%%", lpct*100)
pct_df <- data.frame(
  name = rownames(lcount), 
  count = lcount, 
  pct = lpct, label = pctlab, 
  use.name = FALSE)

# Plot a bar chart of percentage used for each digit from the pct_df source.
ghisto <- ggplot(data = pct_df, mapping = aes(x = name, y = pct, label = label)) +
  geom_col(fill = "goldenrod") + 
  geom_text(vjust = 2) +
  scale_y_continuous(name = "Percentage", labels = scales::percent) +
  labs(
    title = "Distribution of Digit Representations Across 1400 Samples",
    x = "Handwritten Digit",
    fill = "Digit")

# Display the plot.
ghisto

```

## Create a dotplot of digit spreads across all observations

```{r visual_eda_digit_spread_dotplot}

##-------------------------------------------------------------------------------------------------

#' create_spread_plot - Generate a dot plot of the distribution of each digit across the input
#'   data set.
#'
#' @param x Input data set assumed to have a field named "label"
#' @param subtitle Subtitle for the plot to be returned.
#'
#' @return A dot plot of the digit distributions in the input data frame.
#' @export
#'
#' @examples create_spread_plot(train_num, "Original distribution")
create_spread_plot <- function (x, subtitle) {
  # Create a helper dataframe consisting of the data index (row) as an integer and the
  # data value (digit label) as a character string.
  spread_df <- data.frame(
    index = as.integer(1:nrow(x)), 
    value = as.character(x[, "label"]), 
    stringsAsFactors = FALSE)
  # Create a dotplot from the spread_df helper dataframe.
  gdot <- ggplot(data = spread_df, mapping = aes(x = value, y = index)) +
    geom_point(size = 1) +
    scale_y_continuous(name = "Observation", breaks = seq(from = 0, to = nrow(x), by = 100)) +
    labs(
      title = "Spread of Each Digit Across All Observations",
      subtitle = subtitle,
      x = "Digit")
  # Return the generated dotplot
  return(gdot)
}

# Plot the distribution of the original data.
gdot_original <- create_spread_plot(train_num, "Original distribution")
gdot_original

# Now try shuffling and replotting
train_num_shuffle1 <- train_num[sample(nrow(train_num)), ]
gdot_shuffle1 <- create_spread_plot(train_num_shuffle1, "Suffled Distribution #1")
gdot_shuffle1

# Shuffle and replot one more time.
train_num_shuffle2 <- train_num_shuffle1[sample(nrow(train_num_shuffle1)), ]
gdot_shuffle2 <- create_spread_plot(train_num_shuffle2, "Suffled Distribution #2")
gdot_shuffle2

# Create a 1x2 grid display of the two shuffled plots.
grid.arrange(gdot_shuffle1, gdot_shuffle2, nrow = 1)
```

## Display observation images

```{r display_an_image}

##-------------------------------------------------------------------------------------------------

#' display_image - Display a digitized version of the hand-drawn image from a single observation.
#'
#' @param x Image record consisting of a label field and 784 pixel fields. 
#'   This function requires that the pixel fields are numerics.
#'
#' @return A ggplot of the digitized image.
#' @export
#'
#' @examples display_image(train_num[123,])
display_image <- function (x) {
  # Define error messages
   emsg_recordlen <- "Incorrect record length sent to display_image. Expected 785 values."
   emsg_vartype <- "Incorrect data types sent to display_image. Numeric pixel values expected."
   
  # Verify that a record of the proper length was input
  assert_that(ncol(x) == 785, msg = emsg_recordlen)
  # Verify that the pixel fields are all numeric.
  assert_that(all(apply(X = x[, -1], MARGIN = 2, FUN=is.number)), msg = emsg_vartype)
  
  rownums <- as.vector(sapply(X=c(1:28), FUN= function(x) rep(x, times=28)))
  colnums <- rep(c(1:28), times = 28)
  df <- data.frame(drow = rownums, dcol = colnums, ddat = unlist(x[2:785]), row.names = NULL)
  
  g <- ggplot(data = df, mapping = aes(x = dcol, y = -drow, fill = ddat)) +
    geom_tile() +
    scale_fill_continuous(low = "white", high = "black") +
    coord_fixed(ratio = 4/3) + 
    theme(
      legend.position = "none", 
      axis.text.x=element_blank(), axis.ticks.x=element_blank(), 
      axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
    labs(x = paste("label = ", x$label, sep=""))
  
  return(g)
}

set.seed(random_seed)

# Create a list of n random images.
ilist <- list()
num_images <- 28
sample_value <- round(runif(n = num_images, min=1, max = 1400))
for (i in 1:num_images) {
  # Create an image graphic
  ilist[[i]] <- display_image(train_num[sample_value[i],])
}

# Display a single digit.
ilist[[16]]

# Display all created images in a grid.
g <- grid.arrange(
  ilist[[1]], ilist[[2]], ilist[[3]], ilist[[4]], 
  ilist[[5]], ilist[[6]], ilist[[7]], ilist[[8]],
  ilist[[9]], ilist[[10]], ilist[[11]], ilist[[12]], 
  ilist[[13]], ilist[[14]], ilist[[15]], ilist[[16]],
  ilist[[17]], ilist[[18]], ilist[[19]], ilist[[20]],
  ilist[[21]], ilist[[22]], ilist[[23]], ilist[[24]],
  ilist[[25]], ilist[[26]], ilist[[27]], ilist[[28]],
  nrow = 4)


```
# RESULT VISUALIZATION FUNCTIONS

```{r plot_results}

##-------------------------------------------------------------------------------------------------
## RESULTS VISUALIZATION FUNCTIONS

# FUNCTION: digit_model_heatmap
#  PURPOSE: Generate a heatmap showing the success of a model at predicting digitized handwritten 
#           images. Geneally this function is called with the output of a single fold. Therefore, 
#           the reported percent success rate may not be equal to the comp
#    PARAM: list_of_cats - A list of confusion matrices from all the k-folds runs of a model.
#    PARAM: modelname - The name of the model that was executed as a character string. 
#    PARAM: paramstr - A character string describing the specific parameters for the model.
#  RETURNS: A ggplot of the success heatmap that was generated.

digit_model_heatmap <- function (list_of_cmats, modelname, paramstr) {
  
  # Compute a single confusion matrix for the total results
  total_results <- list_of_cmats[[1]]
  for (i in 2:k_folds) {
    total_results <- total_results + list_of_cmats[[i]] 
  }

  # Create desired axis labels and label break points/
  breaks1_10 <- c(0:9)
  labels0_9  <- as.character(c(0:9))
  
  # Turn the confusion matrix into a long data frame.
  verify_df <- melt(total_results)
  names(verify_df) <- c("pred", "label", "freq")
  
  # Compute the overall success rate.
  success_rate <- sum(diag(total_results))/sum(total_results) * 100
  
  # Create the heatmap plot
  g <- ggplot(data = verify_df, mapping = aes(x = pred, y = label, fill = freq, label = freq)) +
    geom_tile(color = "gray50") +
    geom_text() +
    scale_fill_gradient(low = "white", high = "navy") + 
    scale_y_continuous(expand=c(0,0), breaks = breaks1_10, labels = labels0_9) +
    scale_x_continuous(expand=c(0,0), breaks = breaks1_10, labels = labels0_9) +
    labs(
      title = sprintf("Success Map for Digit Recognition Using Model %s", modelname),
      subtitle = sprintf("Parameter: %s\nSuccess rate = %5.02f%%", paramstr, success_rate),
      x = "Predicted Value", 
      y = "Test Label",
      fill = "Volume")
  
  # Return the result
  return(g)
}
```

# PERFORM MODEL-BASED ANALYSES

## K-Nearest Neighbor Analysis

Results for various k
K | Result %
- | ----------
1 | 0.8914286
3 | 0.8864286
5 | 0.8921429
7 | 0.8850000
9 | 0.8864286

```{r run_k_nearest_neghbor}

##-------------------------------------------------------------------------------------------------
## K-Nearest Neighbor Analysis
#
# Results for various k
#  K | Result %
#  - | ----------
#  1 | 0.8914286
#  3 | 0.8864286
#  5 | 0.8921429 ***
#  7 | 0.8850000
#  9 | 0.8864286

cat("Analysis #1: KNN.\n")

# Set the random seed and assign the current k vaue.
set.seed(random_seed)
k <- 5

# Initialize data structures to hold results.
success <- vector(mode = "numeric")
results <- list()

# Perform k-folds testing
for (i in 1:k_folds) {
  # Define the range of indices to be held out for cross-validation
  test_range <- c(floor(fold_size * (i - 1) + 1):floor(fold_size * i))
  knn_train <- train_num[-test_range, -1]
  knn_test <- train_num[test_range, -1]
  knn_train_labels <- train_num$label[-test_range]
  knn_test_labels <- train_num$label[test_range]
  
  # Run the KNN model. The model is tested in the same call in which it is trained. 
  knn_model <- class::knn (train=knn_train, test=knn_test, cl=knn_train_labels, k=k, prob=TRUE)
  # print(knn_model)
  
  # Generate, print and record the confusion matrix for the fold.
  t <- (table(knn_model, knn_test_labels))
  print(t)
  results[[i]] <- t
  
  # Compute and save the success ratio
  success <- c(success, sum(diag(t))/sum(t))
  msg <- sprintf("Test %d: Percent predicted %5.1f%%\n", i, success[i]*100)
  cat(msg, sep="")
}

# Report on the overall success percentage
overall_success <- sum(success)/NROW(success)
cat("Overall: ", round(overall_success*100, digits = 1), "% correctly predicted.\n", sep="")

# Create the success map.
ghm_knn <- digit_model_heatmap(results, "KNN", "k=5")
ghm_knn

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
# Generate a graphic plotting all KNN results.

# Record all models executed and their results.
model_variations <- data.frame(
  k = c(1, 3, 5, 7, 9),
  resultpct = c(0.8914286, 0.8864286, 0.8921429, 0.885, 0.8864286))

# Create a dot & line plot of the results
g_knn <- 
  ggplot(
    data=model_variations,
    mapping=aes(x = k, y = resultpct, label = round(resultpct*100, 2) %s+% "%")) +
  geom_point() +
  geom_line(color = "gray", linetype = "dotted") +
  geom_text(vjust = -2, size = 3) +
  scale_x_continuous(breaks = model_variations$k) + 
  scale_y_continuous(name = "Percent Correct", labels = scales::percent, limits = c(0.885, 0.895)) +
  labs(
    title = "Results of KNN Analysis",
    subtitle = "Only variation was to K")
g_knn

```

```{r clean_knn}

rm(k, success, results, overall_success)
rm(knn_train, knn_test, knn_train_labels, knn_test_labels, t)
rm(test_range, knn_model)

```

## SVM Analysis

Results for various kernel, costs
Kernel     | Data    | Cost   | Coef0 | Result %
---------- | ------- | ------ | ----- | ---------
polynomial | binary  |    1.0 |     0 | 11.929%
polynomial | binary  |    1.0 |  6000 | 11.929%
polynomial | binary  |  100.0 |     0 | 82.286%
polynomial | binary  |  250.0 |     0 | 91.143%
polynomial | binary  |  500.0 |     0 | 91.357%
polynomial | binary  | 1000.0 |     0 | 91.357%
polynomial | numeric |    0.1 | -     | 90.643%
polynomial | numeric |    0.1 |  6000 | 92.071%
polynomial | numeric |    0.5 | -     | 90.643%
polynomial | numeric |    1.0 | -     | 90.643%
polynomial | numeric |    5.0 | -     | 90.643%
polynomial | numeric |   25.0 |   100 | 90.857%
polynomial | numeric |  250.0 |    50 | 90.643%
polynomial | numeric |  250.0 |   100 | 90.857%
polynomial | numeric |  250.0 |   250 | 90.857%
polynomial | numeric |  250.0 |  1000 | 91.429%
polynomial | numeric |  250.0 |  2000 | 91.929%
polynomial | numeric |  250.0 |  5000 | 91.929%
polynomial | numeric |  250.0 |  5500 | 91.929%
polynomial | numeric |  250.0 |  6000 | 92.071%
polynomial | numeric |  250.0 |  7000 | 91.929%
polynomial | numeric |  250.0 |  7500 | 91.571%
polynomial | numeric |  250.0 | 10000 | 91.214% 
polynomial | numeric | 2500.0 |  6000 | 92.071%
linear     | numeric |    0.1 | -     | 88.786%
linear     | numeric |    1.0 | -     | 88.786%
linear     | numeric |  100.0 | -     | 88.786%
linear     | numeric | 1000.0 | -     | 88.786%
linear     | numeric | 9000.0 | -     | 88.786%
sigmoid    | numeric |    0.1 |     0 | 11.929%
sigmoid    | numeric |    1.0 |     0 | 11.929%
sigmoid    | numeric |    1.0 |  1000 | 11.929%
sigmoid    | numeric |    1.0 |  5000 | 11.929%
sigmoid    | numeric |    1.0 | 10000 | 11.929%
sigmoid    | numeric |   10.0 |     0 | 11.929%
sigmoid    | numeric |  100.0 |     0 | 11.929%
sigmoid    | numeric | 9000.0 |     0 | 11.929%
sigmoid    | binary  |    1.0 |     0 | 85.143%
sigmoid    | binary  |   10.0 |     0 | 89.143%
sigmoid    | binary  |   15.0 |     0 | 89.643%
sigmoid    | binary  |   16.0 |     0 | 89.857%
sigmoid    | binary  |   17.0 |     0 | 89.571%
sigmoid    | binary  |   18.0 |     0 | 89.500%
sigmoid    | binary  |   25.0 |     0 | 89.000%
sigmoid    | binary  |   50.0 |     0 | 88.214%
sigmoid    | binary  |  100.0 |     0 | 88.143%
sigmoid    | binary  |  200.0 |     0 | 87.786%
sigmoid    | binary  | 1000.0 |     0 | 87.786%
sigmoid    | binary  | 5000.0 |     0 | 87.786%
sigmoid    | binary  | 9000.0 |     0 | 87.786%

```{r run_svm_model}

##-------------------------------------------------------------------------------------------------
## SVM Analysis
##
## Run SVM 
#
# Results for various kernel, costs
# Kernel     | Data    | Cost   | Coef0 | Result %
# ---------- | ------- | ------ | ----- | ---------
# polynomial | binary  |    1.0 |     0 | 11.929%
# polynomial | binary  |    1.0 |  6000 | 11.929%
# polynomial | binary  |  100.0 |     0 | 82.286%
# polynomial | binary  |  250.0 |     0 | 91.143%
# polynomial | binary  |  500.0 |     0 | 91.357%
# polynomial | binary  | 1000.0 |     0 | 91.357%
# polynomial | numeric |    0.1 |     0 | 90.643%
# polynomial | numeric |    0.5 |     0 | 90.643%
# polynomial | numeric |    1.0 |     0 | 90.643%
# polynomial | numeric |    1.0 |  6000 | 92.071%
# polynomial | numeric |    5.0 | -     | 90.643%
# polynomial | numeric |   25.0 |   100 | 90.857%
# polynomial | numeric |  250.0 |    50 | 90.643%
# polynomial | numeric |  250.0 |   100 | 90.857%
# polynomial | numeric |  250.0 |   250 | 90.857%
# polynomial | numeric |  250.0 |  1000 | 91.429%
# polynomial | numeric |  250.0 |  2000 | 91.929%
# polynomial | numeric |  250.0 |  5000 | 91.929%
# polynomial | numeric |  250.0 |  5500 | 91.929%
# polynomial | numeric |  250.0 |  6000 | 92.071%
# polynomial | numeric |  250.0 |  7000 | 91.929%
# polynomial | numeric |  250.0 |  7500 | 91.571%
# polynomial | numeric |  250.0 | 10000 | 91.214% 
# polynomial | numeric | 2500.0 |  6000 | 92.071%
# linear     | numeric |    0.1 | -     | 88.786%
# linear     | numeric |    1.0 | -     | 88.786%
# linear     | numeric |  100.0 | -     | 88.786%
# linear     | numeric | 1000.0 | -     | 88.786%
# linear     | numeric | 9000.0 | -     | 88.786%
# sigmoid    | numeric |    0.1 |     0 | 11.929%
# sigmoid    | numeric |    1.0 |     0 | 11.929%
# sigmoid    | numeric |    1.0 |  1000 | 11.929%
# sigmoid    | numeric |    1.0 |  5000 | 11.929%
# sigmoid    | numeric |    1.0 | 10000 | 11.929%
# sigmoid    | numeric |   10.0 |     0 | 11.929%
# sigmoid    | numeric |  100.0 |     0 | 11.929%
# sigmoid    | numeric | 9000.0 |     0 | 11.929%
# sigmoid    | binary  |    1.0 |     0 | 85.143%
# sigmoid    | binary  |   10.0 |     0 | 89.143%
# sigmoid    | binary  |   15.0 |     0 | 89.643%
# sigmoid    | binary  |   16.0 |     0 | 89.857%
# sigmoid    | binary  |   17.0 |     0 | 89.571%
# sigmoid    | binary  |   18.0 |     0 | 89.500%
# sigmoid    | binary  |   25.0 |     0 | 89.000%
# sigmoid    | binary  |   50.0 |     0 | 88.214%
# sigmoid    | binary  |  100.0 |     0 | 88.143%
# sigmoid    | binary  |  200.0 |     0 | 87.786%
# sigmoid    | binary  | 1000.0 |     0 | 87.786%
# sigmoid    | binary  | 5000.0 |     0 | 87.786%
# sigmoid    | binary  | 9000.0 |     0 | 87.786%

cat("Analysis #2: SVM\n")

# Set the radom seed and define 3 of the 4 key parameters. The fourth parameter is set by altering
# "train_num" to/from "train_binary" in three successive lines of the following code. 
set.seed(random_seed)
kernel <- "polynomial"
c <- 1
coef0 <- 6000

# Initialize result reoprting data structures.
success <- vector(mode = "numeric")
results <- list()

# Perform k-folds analysis
for (i in 1:k_folds) {
  # Define the range of indices to be held out for cross-validation
  test_range <- c(floor(fold_size * (i - 1) + 1):floor(fold_size * i))
  train <- train_num[-test_range, ]
  test <- train_num[test_range, ]
  test_labels <- train_num$label[test_range]
  
  # Generate the SVM model
  svm_model <- e1071::svm(
    formula = label~., data = train, kernel = kernel, cost = c, coef0 = coef0, scale = FALSE)
  
  # Perform a prediction with the test data
  svm_pred <- predict(svm_model, test, type = "class")
  
  # Generate, print and record the confusion matrix
  t <- (table(svm_pred, test_labels))
  print(t)
  results[[i]] <- t
  
  # Compute and save the success ratio
  success <- c(success, sum(diag(t))/sum(t))
  msg <- sprintf("Test %d: Percent predicted %5.1f%%\n", i, success[i]*100)
  cat(msg, sep="")
}

# Report on the overall success percentage
overall_success <- sum(success)/NROW(success)
cat("Overall: ", round(overall_success*100, digits = 3), "% correctly predicted.\n", sep="")

# Create the success map.
ghm_svm<- digit_model_heatmap(
  results, "SVM", "decimal numerics; kernel=polynomial; c=1, coef0=6000")
ghm_svm



```

### Analyze SVM results

```{r analyze_svm_results}

##-------------------------------------------------------------------------------------------------
## GENEREATE SVM RESULTS SCATTER PLOT REPORTS

svm_results_filepath <- file.path(cwd, data_subdir, "svm_test_results.csv")
assertthat::assert_that(file.exists(svm_results_filepath))

# Read SVM experiment results
svm_results <- read.csv(file = svm_results_filepath, header = TRUE)

# Change all NA values of Coef0 to 0 - Coef0's default value.
svm_results$Coef0[is.na(svm_results$Coef0)] <- 0

# Get the max values
best_values <- svm_results[svm_results$Result == max(svm_results$Result), ]

# Create a base plot of all points.
g_res_all <- ggplot(data = svm_results, mapping = aes(x = Cost, y = Result, shape = Kernel, color = Data, size = Coef0)) +
  geom_point() +
  scale_x_log10() +
  scale_shape_manual(values=c(3, 4, 1)) +
  scale_color_brewer(palette = "Set1") +
  labs(
    title = "Results of 50 SVM Tests for Handwritten Digit Recognition")
g_res_all

# Create a second plot by focusing only on the top portion of the base plot.
g_res_top <- g_res_all + ylim(80, NA)  +
  geom_label(
    data = best_values, 
    mapping = aes(label = Result %s+% "%"), 
    size = 2.5, color = "black", nudge_x = 0.3)
g_res_top


```

```{r remove_vars2}

# Remove no-longer-needed variables.
rm(kernel, c, coef0)
rm(success, results, overall_success)
rm(train, test, test_labels, t, test_range)
rm(svm_pred)
rm (msg)
rm(svm_results_filepath, svm_results)

```

## Random Forest Analysis

Results

Data    | Trees | Result
------- | ----- | ------
numeric |   500 | 91.500%
numeric |   350 | 91.571%
numeric |   340 | 91.143%
binary  |   500 | 91.857%
binary  |   430 | 92.000%
binary  |   385 | 91.786%

```{r random_forest_analysis}

##-------------------------------------------------------------------------------------------------
## Random Forest Analysis
##
## The analysis personally preferred by Forrest Gump. Bubba, on the other hand, has a liking
## for SVMs. Who knew? 
#

cat("Analysis #3: Random Forest\n")

# Set the random seed and the ntrees parameter. The remaining parameter is set by altering four
# lines of the following code to/from train_num and train_binary.
set.seed(random_seed)
ntrees <- 430

# Initialize results data structures.
success <- vector(mode = "numeric")
results <- list()

# Perform k-folds analysis...
for (i in 1:k_folds) {
  # Define the range of indices to be held out for cross-validation
  test_range <- c(floor(fold_size * (i - 1) + 1):floor(fold_size * i))
  train <- train_binary[-test_range, -1]
  test <- train_binary[test_range, -1]
  train_labels <- train_binary$label[-test_range]
  test_labels <- train_binary$label[test_range]
  
  # Run the random forest model. (Time the model as it is running to get a feel for the execution
  # performance of the model.) Testing is performed along with creation of the model so there is no
  # need for any subsequent calls to predict(). 
  print(system.time(
    rf_model <- randomForest::randomForest(
      x = train, y = train_labels, 
      xtest = test, ytest = test_labels, 
      keep.forest = TRUE, ntree = ntrees)))
  
  # Create, print and record the confusion matrix
  t <- rf_model$test$confusion
  print(t)
  results[[i]] <- t

  # Compute and save the success ratio
  success <- c(success, sum(diag(t[, -11]))/sum(t[, -11]))
  msg <- sprintf("Test %d: Percent predicted %5.1f%%\n", i, success[i]*100)
  cat(msg, sep="")
}

# Plot the error rates vs. number of trees
plot(rf_model, main = "Random Forest Error Rates vs. Tree Growth")

# Report on the overall success percentage
overall_success <- sum(success)/NROW(success)
cat("Overall: ", round(overall_success*100, digits = 3), "% correctly predicted.\n", sep="")

# Create the success map.
key_results <- lapply(X=results, FUN = function(x) x[1:10,1:10])
ghm_rf<- digit_model_heatmap(
  key_results, "Random Forest", "binary data; ntrees = 385")
ghm_rf

#- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
# Plot the results of random forest

# Create a data frame to hold parameters and results.
model_variations <- data.frame(
  datatype  = c("decimal", "decimal", "decimal", "binary", "binary", "binary"),
  trees     = c(500, 350, 340, 500, 430, 385),
  resultpct = c(0.915, 0.91571, 0.9143, 0.91857, 0.92, 0.91786))

# Create a scatter plot of the results.
g_rf <- 
  ggplot(
    data = model_variations,
    mapping = aes(
      x=trees, y=resultpct, 
      label=round(resultpct*100, 2) %s+% "%", 
      group=datatype, color=datatype, shape=datatype)) +
  geom_point() +
  geom_line(linetype = "dotted") +
  geom_text(vjust = -2, size = 3) +
  scale_color_brewer(palette = "Set1") +
  scale_y_continuous(name = "Percent Correct", labels = scales::percent, limits = c(0.914, 0.921)) +
  labs(
    title = "Results of Random Forest Analysis",
    subtitle = "Binary data models in red. Decimal data models in blue.",
    x = "Maximum Number of Trees") +
  theme(legend.title = element_blank())
g_rf

```

# FINAL RUN OF THE WINING ALGORITHM

```{r run_best_on_test_data, eval=FALSE}

##-------------------------------------------------------------------------------------------------
## FINALE - RUN THE WINNING ALGORITHM ACROSS THE PROVIDED TEST DATA
##
## Run an Naive Bayes analysis using binary factor pixel values on the test data.

cat("FINALE: Pixel data as binary factors on supplied test data\n")

# Perform a prediction with the test data
svm_final_pred <- predict(svm_model, test_num, type = "class")
final_df <- data.frame(digit = svm_final_pred, stringsAsFactors = FALSE)
ghisto_final <- ggplot(data <- final_df, mapping = aes(x = digit)) +
  stat_count(bins = 10, color="white", fill = "seagreen") +
 # scale_x_discrete(name = "Digit Predicted", breaks = seq(from = 0, to = 9, by = 1)) +
  labs(
    title = "Prediction Volumes for the 999-element Test Data Set",
    subtitle = "Using SVM with coef0 = 6000, c = 1",
    y = "Prediction Volume")
ghisto_final

```


